%!TEX root = ../Thesis.tex
\chapter*{\englishprefacename}
\addcontentsline{toc}{chapter}{\englishprefacename}

This master's thesis is a part of the requirements for the master degree at the Department for Engineering Cybernetics at the Norwegian University of Science and Technology (NTNU). The thesis was done under the supervision of Associate professor Fredrik Dessen and in collaboration with SINTEF Energy Research under the additional supervision of Dr Elisa Magnelli, and Dr Michael Becidan. 

The thesis is not based on a previous specialization project. However it is based on previous work at SINTEF Energy Research, as well as a report  of a previous summer-student, Herman Øie Kolden. As with all of the reports from the summer-students at SINTEF, the report by \cite{summer_student} was never published. \todo[inline]{Check if the report was ever published}

\noindent
SINTEF provided am in-house simulator developed during the Waste-to-Energy 2030 project, co-funded by industry and public partners and the Research Council of Norway under EnergiX program (WtW, 2030, 280949).

\noindent
The model was expanded to include an implementation of an A/B-controller and a cascaded A/B controller through the work done in \cite{summer_student}. This model was used in chapter \ref{cha:experimental_results}. A Model Predictive Controller had also been developed using one of MATLAB's toolboxes, but due to the relatively long sampling-time, it had no way to keep up with the performance of the A/B-controller. The implementation of the Eigensystem Realisation Algorithm was based on the video-lectures on Youtube by Steve Brunton. All plots were made using MATLAB's built-in functions. The same goes for most complex mathematical operations that have a build-in MATLAB-function, like singular value decomposition, and conversion between continuous and discrete Linear Time-Invariant systems. 


\noindent
I would especially like to thank one of my supervisors, Elisa, for taking time out of her day, helping me with the mysterious problems that resulted from the simulator. Also, I'd like to thank Fredrik, as well, for answering most of my control-related questions and for advising me to take more time, so that I could deliver a more functioning product. Finally, I'd like to thank my family for giving me the time, space and support needed for completing this project. 

\hfill {\it Jørn Bøni Hofstad}\\[-6mm]

\hfill {\it Trondheim, August 2020}
%

\clearpage
